Best Practices for Single Alchemical Free Energy Calculations [Article v\versionnumber]

### 2025-10-02 paper/manuscript.tex
Do not require ligands for ABFE have the same binding pose in the checklist section
Pointed out by @sboresch

MD trajectories are extremely sensitive to their starting conditions and hence one-off simulations can never be reproducible~\cite{coveney2016, MDbook2025}. Thus, it is recommended to perform ensemble simulations in order to perform ensemble averaging to obtain thermodynamic quantities in a reliable manner. Bhati \textit{et al.} developed the TIES method~\cite{bhati2017, bhati2022, bhati2025} which is based on the use of ensemble methods, where the ensemble size can be adjusted based on available computational resource and required precision~\cite{wan2023}. Studies have shown that, in general, an ensemble size of 5 and 10 members is a reasonable starting point for alchemical RBFE and ABFE calculations, respectively, but recommendations are provided to adjust them based on some rules of thumb~\cite{bhati2022, bhati2025}. 
The sample standard deviation will give a crude estimate of the reliability of the estimates, and whether the precision is sufficient for the problem at hand.

@@

\begin{checklist}{Running absolute simulations}
        \textbf{Steps to running your absolute alchemical free energy calculations}
\begin{itemize}
 \item Check your ligands have the same, biologically correct binding pose
        \item Make sure your \textlambda-scheduling is  appropriate
        \item Check if your ligands are discharging and decoupling correctly
        \item Set up your restraints correctly
        \item Subsample the data to estimate the variance of your free energy estimation protocol (but not to estimate the values themselves)
        \item Apply the appropriate correction terms
        \item Ensure sufficient configurational sampling for slow degrees of freedom (e.g., check for kinetic trapping in metastable states).
\end{itemize}
\end{checklist}

>>>>

MD trajectories are extremely sensitive to their starting conditions and hence one-off simulations can never be reproducible~\cite{coveney2016, MDbook2025, caves_locally_1998}. Thus, it is recommended to perform ensemble simulations in order to perform ensemble averaging to obtain thermodynamic quantities in a reliable manner. Bhati \textit{et al.} developed the TIES method~\cite{bhati2017, bhati2022, bhati2025} which is based on the use of ensemble methods, where the ensemble size can be adjusted based on available computational resource and required precision~\cite{wan2023}. Studies have shown that, in general, an ensemble size of 5 and 10 members is a reasonable starting point for alchemical RBFE and ABFE calculations, respectively, but recommendations are provided to adjust them based on some rules of thumb~\cite{bhati2022, bhati2025}. 
The sample standard deviation will give a crude estimate of the reliability of the estimates, and whether the precision is sufficient for the problem at hand.

@@

\begin{checklist}{Running absolute simulations}
        \textbf{Steps to running your absolute alchemical free energy calculations}
\begin{itemize}
 \item Check your ligands have the biologically correct binding pose(s)
        \item Make sure your \textlambda-scheduling is  appropriate
        \item Check if your ligands are discharging and decoupling correctly
        \item Set up your restraints correctly
        \item Subsample the data to estimate the variance of your free energy estimation protocol (but not to estimate the values themselves)
        \item Apply the appropriate correction terms
        \item Ensure sufficient configurational sampling for slow degrees of freedom (e.g., check for kinetic trapping in metastable states).
\end{itemize}
\end{checklist}
